---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

* Unzipping the zip data file, reading the resulting data file into "df", picking out only the rows without NA's of "df" and saving them in the data frame "df_REDUCE".

```{r}
unzip("activity.zip")
df <- read.csv(unzip("activity.zip",list = TRUE)[1,1])
df_REDUCE <- df[!is.na(df$steps),]

```

## What is mean total number of steps taken per day?

* Calculating the total, average, and median values of number of steps taken every single day, and group them in one single data frame
```{r, echo = "TRUE"}
LVS <- levels(df_REDUCE$date) # Stocking all the different dates

MEDIAN <- AVERAGE <- TOTAL_STEPS <- vector(mode = "numeric", length = length(LVS)) 
# Creating vectors for stocking the median, average and total of the number of steps of each day. Note that these vectors are of the same length of that of LVS, the vector storing distinct dates

for(i in 1:length(LVS)) { 
  TOTAL_STEPS[i] = sum(df_REDUCE[df_REDUCE$date == LVS[i],]$steps, na.rm = TRUE) 
  AVERAGE[i] = mean(df_REDUCE[df_REDUCE$date == LVS[i],]$steps,na.rm = TRUE) 
  MEDIAN[i] = median(df_REDUCE[df_REDUCE$date == LVS[i],]$steps, na.rm = TRUE) 
} # Calculating the total, average, and median values of number of steps taken every single day

df_daily_statistics <- data.frame(Dates = as.Date(LVS, "%Y-%m-%d"), Total_steps = TOTAL_STEPS, Average = AVERAGE, Median = MEDIAN) # Group these results into a single data frame
```

* Plotting histogram of daily total number of steps

```{r}
cutting.points = seq(0, max(df_daily_statistics$Total_steps), length.out = 10)
hist(df_daily_statistics$Total_steps, col = "lightblue", xlab = "total number of steps each day", main = "Histogram of daily total number of steps",breaks= cutting.points)
```


* Reporting the results of the total, average, and median of the number of steps each day

```{r}
print(df_daily_statistics)
```

## What is the average daily activity pattern?

* Computing the average numbers of steps in specific 5-min intervals, averaged over all available days
```{r}

INTERVALS <- seq(0,max(df$interval,na.rm = TRUE), by = 5)
Ave_interval_steps <- vector(mode = "numeric", length = length(INTERVALS))
for(i in 1:length(INTERVALS)) {
  Ave_interval_steps[i] <- mean( df$steps[df$interval == INTERVALS[i]] , na.rm = TRUE)
}
```

* Plotting the the average numbers of steps in specific 5-min intervals, against the intervals
```{r}
plot(INTERVALS,Ave_interval_steps,type ="l",xlab = "5-minute intervals", ylab = "Average number of steps")
```


* Reporting the interval with maximum average number of steps

```{r}
MAX <- round( max(Ave_interval_steps, na.rm = TRUE),4) 
MAX_ind <- which.max(Ave_interval_steps)
message("The 5-minute interval with maximum average number of steps is [",MAX_ind*5-5,",",MAX_ind*5,").", " with the maximum value ",MAX,"." , sep="")
```


## Imputing missing values

* Reporting the number of observations with missing value of "step"

```{r}
NAS <- is.na(df$steps) # Create a vector marking the location of NA's, where the indices of TRUE's are the indices of the rows with missing step numbers (step = NA)
sum(NAS) # Obtaining total number of TRUE's, ie, the total number of rows with NA in the data frame
```

* Filling NA's with the average value of that 5-minute interval across all days, and save the result in a new data frame named df_noNA

```{r}
STP_NoNA <- df$steps

for(i in 1:length(STP_NoNA)){
  if(is.na(STP_NoNA[i]) & !is.nan(Ave_interval_steps[df$interval[i]/5+1] )) STP_NoNA[i] <- Ave_interval_steps[df$interval[i]/5+1]
}

df_NoNA <- data.frame( steps = STP_NoNA, df[,2:3] )
```

* Computing the total, average, and median values of the number of steps of each day using the data frame with missing data filled in

```{r}
MEDIAN2 <- AVERAGE2 <- TOTAL_STEPS2 <- vector(mode = "numeric", length = length(LVS))

for(i in 1:length(LVS)) { 
  TOTAL_STEPS2[i] = sum(df_NoNA[df_NoNA$date == LVS[i],]$steps) 
  AVERAGE2[i] = mean(df_NoNA[df_NoNA$date == LVS[i],]$steps) 
  MEDIAN2[i] = median(df_NoNA[df_NoNA$date == LVS[i],]$steps) 
} # Calculating the total, average, and median values of number of steps taken every single day
```

* Plot the histogram with filled-in missing values

```{r}
hist(TOTAL_STEPS2, col = "blue", xlab = "total number of steps each day", main = "Histogram of daily total number of steps",breaks= cutting.points)
```


* Organizing the results obtained with filled-in missing values into a data frame, and report the result
```{r}
df_daily_statistics2 <- data.frame(Dates = as.Date(LVS, "%Y-%m-%d"), Total_steps = TOTAL_STEPS2, Average = AVERAGE2, Median = MEDIAN2) # Group these results into a single data frame

df_daily_statistics2
```

## Are there differences in activity patterns between weekdays and weekends?

* Creating the vector indicating whether each row is of Weekday or Weekend, and binding this vector with the data frame "df_NoNA", naming the new column "WD.or.WE"

```{r}
Days <- weekdays(as.Date(df_NoNA$date, "%Y-%m-%d"))

for(i in 1:length(Days)) {
  if(Days[i]=="Samedi" | Days[i]=="Dimanche") Days[i] <- "Weekend"
  else Days[i] <- "Weekday"
}
df_NoNA <- cbind(df_NoNA, "WD.or.WE"=factor(Days))
```

* Using the "WD.or.WE" column, computing the average number of steps of each 5-minutes intervals across weekdays and weekends, storing the results in "Ave_interval_steps_WD" and "Ave_interval_steps_WE" respectively

```{r}
Ave_interval_steps_WD <- Ave_interval_steps_WE <- vector(mode = "numeric", length = length(INTERVALS))
for(i in 1:length(INTERVALS)) {
  Ave_interval_steps_WD[i] <- mean( df_NoNA$steps[(df_NoNA$interval == INTERVALS[i])&(Days=="Weekday")] , na.rm = TRUE)
  Ave_interval_steps_WE[i] <- mean( df_NoNA$steps[(df_NoNA$interval == INTERVALS[i])&(Days=="Weekend")] , na.rm = TRUE)
}
```

* Plotting the results in a 2x1-pannel, weekday result in the upper plot and weekend result in the lower plot 

```{r, fig.height=7}
par(mfrow=c(2,1),mar=c(4,5,2,1))
plot(INTERVALS,Ave_interval_steps_WD,xlab="",ylab="Av. Num. of Steps",main="Weekdays",type="l")
plot(INTERVALS,Ave_interval_steps_WE,xlab="5-minute intervals",ylab="Av. Num. of Steps",main="Weekends",type="l")
```
